Katana: Towards Patching as a Part of the ELF Toolchain

### 2010-03-18 paper1/hotpatch/poformat.tex
slight clarification on versioning issue

%\begin{wrapfigure}{l}{0.4\textwidth}
%{\footnotesize
\begin{figure}[ht]
\begin{center}
\begin{verbatim}
  Section Headers:
    [Nr] Name              Type          
    [ 0]                   NULL          
    [ 1] .strtab           STRTAB        
    [ 2] .symtab           SYMTAB        
    [ 3] .text.new         PROGBITS      
    [ 4] .rodata.new       PROGBITS      
    [ 5] .rela.text.new    RELA          
    [ 6] .debug_info       PROGBITS      
    [ 7] .debug_abbrev     PROGBITS      
    [ 8] .debug_frame      PROGBITS      
    [ 9] .rel.debug_info   REL           
    [10] .rel.debug_frame  REL           
\end{verbatim}
\end{center}
%}
\caption{Headers for the PO}
\label{fig:elfheaders}
%\vspace{-10pt}
%\end{wrapfigure}
\end{figure}
constants to inject. \texttt{.rela.text.new} allows \texttt{.text.new}
to be properly relocated after it is adjusted. While System V based
systems use only relocation sections of type \texttt{SHT\_REL}, we
chose to use \texttt{SHT\_RELA} in our patch objects because they make
addends much easier to keep track of as we relocate from patched
binary to patch object to patched process in memory. This is all
really nothing new, storing ELF sections to be injected in-memory has
been done before in other systems \cite{eresi}. What is new in our patch object is
the inclusion of DWARF sections. The \texttt{.debug\_info} section in
an ordinary executable program contains a tree of DIEs (Debugging
Information Entities) with information about every type, variable, and
procedure in each compilation unit in the program. In a patch object,
we store information only about the procedures and variables which
have changed. This of course includes storing the type information for
changed variables. An example of the DWARF DIE information contained
in a patch can be seen in Figure \ref{fig:dies}.

Note that we store considerably less information about each entity
than is typically contained. This is because we read most of the
information from the DWARF and symbol table information of the
executing process (unless it will have been stripped, then more
information must be stored in the patch). This allows the patch to be
more flexible as it does not require that all variables and procedures
be located at exactly the addresses they were expected to be at when
the patch was generated. This flexibility allows a single patch
between versions \emph{va} and \emph{vb} to patch both an executable
which was originally compiled to \emph{va} and an executable which was
patched from earlier versions to be equivalent to \emph{va}. Ksplice,
one of the few other patchers which operates solely at the binary
level, does not have this capability \cite{ksplice}. Note that patch
versioning is currently a work in progress and not fully implemented.
%\begin{wrapfigure}{r}{0.5\linewidth}
\begin{figure}[ht]
{\small
\begin{center}
\begin{verbatim}
  .debug_info

>>>>

%\begin{wrapfigure}{l}{0.4\textwidth}
%{\footnotesize
\begin{figure}[ht]
\begin{center}
\begin{verbatim}
  Section Headers:
    [Nr] Name              Type          
    [ 0]                   NULL          
    [ 1] .strtab           STRTAB        
    [ 2] .symtab           SYMTAB        
    [ 3] .text.new         PROGBITS      
    [ 4] .unsafe_functions LOUSER+1
    [ 5] .rodata.new       PROGBITS      
    [ 6] .rela.text.new    RELA          
    [ 7] .debug_info       PROGBITS      
    [ 8] .debug_abbrev     PROGBITS      
    [ 9] .debug_frame      PROGBITS      
    [10] .rel.debug_info   REL           
    [11] .rel.debug_frame  REL           
\end{verbatim}
\end{center}
%}
\caption{Headers for the PO}
\label{fig:elfheaders}
%\vspace{-10pt}
%\end{wrapfigure}
\end{figure}
constants to inject. \texttt{.rela.text.new} allows \texttt{.text.new}
to be properly relocated after it is adjusted. While System V based
systems use only relocation sections of type \texttt{SHT\_REL}, we
chose to use \texttt{SHT\_RELA} in our patch objects because they make
addends much easier to keep track of as we relocate from patched
binary to patch object to patched process in memory. This is all
really nothing new, storing ELF sections to be injected in-memory has
been done before in other systems \cite{eresi}. What is new in our patch object is
the inclusion of DWARF sections. The \texttt{.debug\_info} section in
an ordinary executable program contains a tree of DIEs (Debugging
Information Entities) with information about every type, variable, and
procedure in each compilation unit in the program. In a patch object,
we store information only about the procedures and variables which
have changed. This of course includes storing the type information for
changed variables. An example of the DWARF DIE information contained
in a patch can be seen in Figure \ref{fig:dies}.

Note that we store considerably less information about each entity
than is typically contained. This is because we read most of the
information from the DWARF and symbol table information of the
executing process (unless it will have been stripped, then more
information must be stored in the patch). This allows the patch to be
more flexible as it does not require that all variables and procedures
be located at exactly the addresses they were expected to be at when
the patch was generated. This flexibility allows a single patch
between versions \emph{va} and \emph{vb} to patch both an executable
which was originally compiled to \emph{va} and an executable which was
patched from earlier versions to be equivalent to \emph{va}. Ksplice,
one of the few other patchers which operates solely at the binary
level, does not have this capability \cite{ksplice}. We will provide a
mechanism for composing patches, such that a patch from version
\emph{va} to \emph{vb} may be composed with a patch from \emph{vb} to
\emph{vc} to produce a patch from \emph{va} to \emph{vb}. Note that
patch versioning is currently a work in progress and not fully
implemented.
%\begin{wrapfigure}{r}{0.5\linewidth}
\begin{figure}[ht]
{\small
\begin{center}
\begin{verbatim}
  .debug_info
